Marlowe: financial contracts on blockchain\thanks{This work is part of the Cardano project and is supported by IOHK,~\url{https://iohk.io}}

### 2018-07-18 tex/paper.tex
Fix typos and address some reviewer's comments

This paper explores the design of a domain specific language, Marlowe,\footnote{Named after Christopher 
Marlowe, the Elizabethan poet, dramatist and spy, who was born and educated in Canterbury,  \url{en.wikipedia.org/wiki/Christopher_Marlowe}}\footnote{Marlowe is available 
from \url{https://github.com/input-output-hk/scdsl}} targetted at the execution of financial contracts in the style of 
Peyton Jones, Eber and Seward~\cite{PeytonJones:2000} on blockchains.
In doing this, we are required to refine the model of contracts in a number of ways in order to fit with a radically 
different context.

Consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants \haskellinline{1}, \haskellinline{2} and \haskellinline{3}. Participant \haskellinline{1} is 
to pay an amount of money to participant \haskellinline{2} on receipt of goods from her. Participant \haskellinline{1} 
pays the money into escrow controlled by participant \haskellinline{3}. There are two options for the money: if two 
participants agree to \haskellinline{pay} it to \haskellinline{2}, that goes ahead; if, on the other hand, two of the 
participants opt to \haskellinline{refund} the money to \haskellinline{1}, that is done instead. The 
\haskellinline{When} construct fires when its first argument -- a condition -- becomes true; the second argument, a 
\haskellinline{Contract} itself, is then performed, and this makes the payment according to the participants' 
\haskellinline{Choice}.

@@

In the traditional model, enforcement of the contract is the responsibility of the legal system. If participant 
\haskellinline{1} does not pay the money into escrow, or participant \haskellinline{3} chooses to keep it for herself, 
then they can be sued for the money (and probably damages), thus providing both legal and 
financial incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no central 
authority, the contract needs to be enforced \emph{by design}.

This means that we require participants to \emph{commit} money to cover all possible expenditure \emph{in advance of the 
contract executing}. In order to make sure that participants continue to engage with a contract, we ensure urgency by 
imposing \emph{timeouts}: money is committed for a finite period only. We also impose a timeout when  waiting for a participant to make a 
commitment to ensure that the contract does not become stuck even if one of the 
participants stops interacting with it. We discuss this particular example in more detail in Marlowe in Section~\ref{section:example-escrow} below.

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Embedding Marlowe as a Haskell \haskellinline{data} type to extend its expressibility.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of embedded Marlowe contracts. 
\end{itemize}
Having established this model and its semantics we are able to do a number of other things, which we discuss in the paper. We can explore how the language will be implemented on an existing blockchain, such as Cardano, and our model has shown us that we will need to consider how the evolution of a contract interacts with the blockchain, with miners in the case of a `proof of work'-based chain, and with users participating in contract execution. We can also perform analyses of Marlowe contracts, based on its formal semantics. Finally, we can explore the design space for Marlowe, and explore potential alternative designs.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themsleves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of payments by 
a participant. Moreover it is also possible for a participant to input an arbitrary value (which we term a ``choice''). 
The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

In this section, we revisit the escrow example that we discussed briefly in the introduction, and show how we can make 
Marlowe contracts that are easier to write, read, and understand, by embedding them into Haskell code, taking advantage 
of the fact that Marlowe contracts are implemented as Haskell terms.

We used Haskell because it is the language in 
which Marlowe is implemented, but it would be easy to embed Marlowe in any other language. It would only be necessary 
to translate its primitives into a data type in that language. In Meadow we use Fay, which is a subset of 
Haskell (see Section~\ref{section:tool}).

@@

Meadow has been mainly written in Haskell and compiled to JavaScript
partially by the Haste compiler~\cite{Haste} and partially by GHCJS \cite{GHCJS};
Meadow also relies on the Blockly library~\cite{Blockly} for providing a visual
editor for smart contracts written in Marlowe. Embedding support in Meadow
is provided throught a pruned and bundled version of the Fay compiler \cite{Fay},
compiled to JavaScript using GHCJS. All text fields that edit and present Meadow and Fay code are
instances of CodeMirror text editor \cite{CodeMirror}.

Blockly's editor allows the user to visualise and edit
smart-contracts as interlocking blocks that can be dragged and dropped
like pieces of a jigsaw puzzle. % (see Figure~\ref{fig:detail-of-block}).
Meadow also provides functionality to generate syntactically correct
and formatted code (that is displayed in the upper right corner of the application),
and to convert the code back to its blockly representation.

Additionally, Meadow allows the user to use Meadow's Fay embedding from within the browser. % (see Figure~\ref{fig:embedded-demo}). 
Meadow includes a pruned copy of the Fay compiler, bundled with its Prelude module, 
its foreign interface, and a module with the definitions of the Marlowe primitives and a function to pretty-print 
Marlowe contracts. This allows the user to compile and execute, inside the browser, Fay
code that generates and prints a Meadow contract. When the user opens the embedded editor, the current Marlowe contract 
is embedded into a template of a Fay program that prints it. The user is able to modify this code in the left 
panel of the editor and use the advanced functionalities provided by Fay like, for example, bindings, 
list-comprehensions, turing-complete functions definitions, etc.

@@

%\paragraph{Sidechains and wallets}
%
%The implementation may split contract execution between the main chain and a sidechain (or chains). This will allow links from Marlowe to work on securing multiparty sidechain games/contracts (MPCs).

@@

We would like to thank IOHK for supporting us: not only has this led us to work on Marlowe, but we have benefitted hugely from collaboration with colleagues including Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiayias, % Darryl McAdams, 
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler. We are  very grateful to Thomas Arts and colleagues from \ae{}ternity for their convivial discussions on blockchain and contracts in G\"oteborg, and to colleagues from the Universities of Kent and Leicester for their comments too.

>>>>

This paper explores the design of a domain specific language, Marlowe,\footnote{Named after Christopher 
Marlowe, the Elizabethan poet, dramatist and spy, who was born and educated in Canterbury,  
\url{en.wikipedia.org/wiki/Christopher_Marlowe}}\footnote{Marlowe is available from 
\url{https://github.com/input-output-hk/scdsl}} targeted at the execution of financial contracts in the style of 
Peyton Jones, Eber and Seward~\cite{PeytonJones:2000} on blockchains. In doing this, we are required to refine the 
model of contracts in a number of ways in order to fit with a radically different context.

Consider the following example of an ``escrow'' contract so that we can explain the motivation 
more concretely. The aim of this contract, written in functional pseudocode in the style of~\cite{PeytonJones:2000} 
involves three participants \haskellinline{1}, \haskellinline{2} and \haskellinline{3}. Participant \haskellinline{1} is 
to pay an amount of money to participant \haskellinline{2} on receipt of goods from her. Participant \haskellinline{1} 
pays the money into escrow controlled by participant \haskellinline{3}.

There are two options for the money: if two participants agree to \haskellinline{pay} it to \haskellinline{2}, that goes 
ahead; if, on the other hand, two of the participants opt to \haskellinline{refund} the money to \haskellinline{1}, that 
is done instead.

The outer primitive \haskellinline{When} waits until the condition -- its first argument -- 
becomes true; in this case, the condition is the that either two participants choose \haskellinline{refund} or two 
participants choose \haskellinline{pay}. The second argument of the \haskellinline{When} is itself another 
\haskellinline{Contract}, which is performed after the condition of the \haskellinline{When} has been met, and it makes 
the payment if two participants chose \haskellinline{pay}, otherwise it redeems previous money commitments.

@@

We discuss this particular example in more detail in Marlowe in Section~\ref{section:example-escrow} below; but it 
already gives us an example of how traditional contracts are fundamentally different from contracts that are meant to 
be run on top of the blockchain. In the traditional model, enforcement of the contract is the responsibility of the 
legal system. If participant \haskellinline{1} does not pay the money into escrow, or participant \haskellinline{3} 
chooses to keep it for herself, then they can be sued for the money (and probably damages), thus providing both legal 
and financial incentives for compliance. On the other hand, in the decentralised blockchain model, where there is no 
central authority, the contract needs to be enforced \emph{by design}.

This means that we must require participants to \emph{commit} money to cover all possible expenditure \emph{in advance 
of the contract executing}. In order to make sure that participants continue to engage with a contract, we ensure 
urgency by imposing \emph{timeouts}: money is committed for a finite period only. We also impose a timeout when waiting 
for a participant to make a commitment to ensure that the contract does not become stuck even if one of the participants 
stops interacting with it.

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Embedding of Marlowe in Haskell to extend its expressibility; that is: Marlowe is defined as a Haskell 
\haskellinline{data} type so that generating Marlowe contracts from Haskell is straightforward.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of Marlowe contracts 
and embedded Marlowe contracts.
\end{itemize}

Having established this model and its semantics we are able to do a number of other things, which we discuss in the 
paper. We can explore how the language will be implemented on an existing blockchain, such as Cardano, and our model has 
shown us that we will need to consider how the evolution of a contract interacts with the blockchain, with miners in the 
case of a `proof of work'-based chain, and with users participating in contract execution. We can also perform analyses 
of Marlowe contracts, based on its formal semantics. Finally, we can explore the design space for Marlowe, and explore 
potential alternative designs.

@@

On the other hand, at each step there are -- potentially, at least -- a variety of inputs available from the 
participants themselves. These include commitments of currency (or ``cash''), redemption of commitments, and claims of 
payments by a participant. Moreover it is also possible for a participant to input an arbitrary value (which we term a 
``choice''). The particular inputs at a given step are described by a value of type \haskellinline{Input}.

@@

In this section, we revisit the escrow example that we discussed briefly in the introduction, and show how we can make 
Marlowe contracts that are easier to write, read, and understand, by embedding them into Haskell code, that is, taking 
advantage of the fact that Marlowe contracts are implemented as Haskell terms to write Haskell programs that generate 
Marlowe code, instead of writing Marlowe directly.

We used Haskell because it is the language in which Marlowe is implemented, but it would be easy to embed Marlowe in any 
other language. It would only be necessary to translate its primitives into a data type in that language. In Meadow we 
use Fay \cite{Fay}, which is a subset of Haskell (see Section~\ref{section:tool}).

@@

Meadow has been mainly written in Haskell and compiled to JavaScript
partially by the Haste compiler~\cite{Haste} and partially by GHCJS \cite{GHCJS};
Meadow also relies on the Blockly library~\cite{Blockly} for providing a visual
editor for smart contracts written in Marlowe. Embedding support in Meadow
is provided through a pruned and bundled version of the Fay compiler \cite{Fay},
compiled to JavaScript using GHCJS. All text fields that edit and present Meadow and Fay code are
instances of CodeMirror text editor \cite{CodeMirror}.

Blockly's editor allows the user to visualise and edit
smart-contracts as interlocking blocks that can be dragged and dropped
like pieces of a jigsaw puzzle. % (see Figure~\ref{fig:detail-of-block}).
Meadow also provides functionality to generate syntactically correct
and formatted code (that is displayed in the upper right corner of the application),
and to convert the code back to its Blockly representation.

Additionally, Meadow allows the user to use Meadow's Fay embedding from within the browser. % (see Figure~\ref{fig:embedded-demo}).

The reason we chose Fay instead of Haskell for Meadow is a technical one: we wanted Meadow to be run completely inside 
the browser, because that makes it easy to deploy, since the server is only required to send the page and not to 
run any Marlowe-related computation. There exist compilers from Haskell to JavaScript but most of them are not easy to 
bootstrap into JavaScript. Among other reasons, they often rely on the API of the host OS to read and write files.

Fay compiler, on the other hand, is mostly written in pure Haskell, and it is not hard to compile to JavaScript by 
using GHCJS. Nevertheless, it does not include a type checker, it relies in GHC's type-checker; and it also 
tries to read its \haskellinline{Prelude} from disk. We have worked around these issues by disabling type-checking 
within Meadow embedded editor, and by embedding all the required modules as constant strings in a modified version 
of the Fay compiler.

In particular, Meadow includes a pruned copy of the Fay compiler, bundled with its Prelude module, its foreign 
interface, and a module with the definitions of the Marlowe primitives and a function to pretty-print 
Marlowe contracts. This allows the user to compile and execute, inside Meadow, Fay code that generates and prints 
a Marlowe contract. When the user opens the embedded editor, the current Marlowe contract is embedded into a template 
of a Fay program that prints it. The user is able to modify this code in the left panel of the editor and use the 
advanced functionalities provided by Fay like, for example, bindings, list-comprehensions, turing-complete functions 
definitions, etc.

@@

%\paragraph{Sidechains and wallets}
%
%The implementation may split contract execution between the main chain and a sidechain (or chains). This will allow 
%links from Marlowe to work on securing multiparty sidechain games/contracts (MPCs).

@@

We would like to thank IOHK for supporting us: not only has this led us to work on Marlowe, but we have benefited 
hugely from collaboration with colleagues including Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles 
Hoskinson, Aggelos Kiayias, % Darryl McAdams,
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler. We are  very 
grateful to Thomas Arts and colleagues from \ae{}ternity for their convivial discussions on blockchain and contracts in 
G\"oteborg, and to colleagues from the Universities of Kent and Leicester for their comments too.
